---
title: "Statistical assignment 1"
author: "Alhussein Elrayah, Student No: 670037045"
date: "11/02/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open data (10 points)

In this assignment you will work with the individual level data from wave 8 of the Understanding Society survey. First, you need to open the data set. Please complete the code below.

```{r}
library(tidyverse)
Data <- read_tsv('C:/Users/alhus/Desktop/Desktop/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab')
```

Now you have got your data frame stored as Data. 

## Select variables (10 points)

The data for Wave 8 of the Understanding Society were collected in 2016-18. Among other things, people were asked the following question: "Should the United Kingdom remain a member of the European Union or leave the European Union?" In this assignment, we will explore how answers to this question depend on sex and age.

First, you need to select the variables for the analysis. You want to keep the following variables: cross-wave individual identifier (*pidp*), support for the UK remaining or leaving the EU (*h_eumem*), sex (*h_sex_dv*), age (*h_age_dv*), and sample origin (*h_memorig*).

Complete the code below to select those variables from the data frame and save the result.

```{r}
Data <- Data %>%
select(pidp, h_eumem, h_sex_dv, h_age_dv, h_memorig)

## Note: Drop the h from each one above
```


## Filter observations (10 points)

To make nationally representative estimates from the Understanding Society data we would need to use weight coefficients. There are many different types of weight coefficients that can be used depending on the question and the level of analysis (see the User Guide, pp. 65-71). We will not do this in this assignment. However, what we want to do is to keep data from the original Understanding Society sample only (ukhls gb 2009-10), dropping  data for Northern Ireland, the BHPS cohort members and ethnic minority boost samples. This will make data closer to be representative for Great Britain. You need to choose the observations where *h_memorig* has the value of 1.

```{r}
Data <- Data %>%
        filter(h_memorig == 1)
```

## Recode data (20 points)

Let us tabulate the variables for EU support, sex, and age.

```{r}
table(Data$h_eumem)
table(Data$h_sex_dv)
table(Data$h_age_dv)
```

You will see that all these variables are numeric. You can learn what the numeric codes mean by checking the codebook here: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/datafile/indresp/wave/8 .

We want to do the following:

1) Recode the variable for EU support as binary (1 for Remain, 0 for Leave), coding all types of missing values (including refusals and "don't know") as NA.
2) Recode sex into a character vector with the values "male" or "female".
3) Recode age into a variable with the following categories: 16 to 25, 26 to 40, 41 to 55, 56 to 70, over 70.

In all cases, we want to create new variables.

```{r}

## Recoding to create new EU variable

Data %>%
mutate(EU = ifelse(h_eumem == 1, "1", ifelse(h_eumem == 2, "0", NA))) %>%

## Recoding to create new character vector sex variable

mutate(sex = ifelse(h_sex_dv == 1, "Male",ifelse(h_sex_dv == 2, "Female", NA))) %>%

## Recoding to create new categorical age variable

mutate(agegr = case_when(
                 between(h_age_dv, 16, 25) ~ "16 to 25 years",
                 between(h_age_dv, 26, 40) ~ "26 to 40 years",
                 between(h_age_dv, 41, 55) ~ "41 to 55 years",
                 between(h_age_dv, 56, 70) ~ "56 to 70 years",
                 h_age_dv > 70 ~ "Over 70 years"
        )) 

```

## Summarise data (20 points)

Let us **dplyr** to calculate how many people in the sample supported Remain and Leave, both as absolute numbers and percentages.

```{r}

library(dplyr)

Data  %>%
  mutate(EU = ifelse(h_eumem == 1, "1", ifelse(h_eumem == 2, "0", NA))) %>%
group_by(EU) %>%
summarise(n = n())

## Finding the proportions

11118/22957 ## 48.4% of respondents voted Remain

9338/22957 ## 40.7% of respondents voted Leave




        
```


Write a couple of sentences with the interpretation of this result. How this compares with the result of the 2016 referendum? Why?

In the 2016 referednum, 51.9% voted to leave and 48.1% voted to remain. The results of the questionnaire are 
close to the actual result of the referendum in terms of the remain vote but fall short of more than 10
percentage points from the leave vote. It is possible this can be attributed in part to the NA category. Also,
leave voters may live in areas which are harder to reach (more rural) for those conducting the survey for more urban-dwelling remian voters.

(Source: https://www.bbc.co.uk/news/politics/eu_referendum/results)

## Summarise data by sex and age (30 points)

Now let us look at the support for Leave and Remain by sex and age. Use your newly created variables.


```{r}

## For Sex

Data  %>%
  mutate(EU = ifelse(h_eumem == 1, "1", ifelse(h_eumem == 2, "0", NA))) %>%
  mutate(sex = ifelse(h_sex_dv == 1, "Male",ifelse(h_sex_dv == 2, "Female", NA)))%>%
group_by(EU, sex) %>%
summarise(n = n())

## Finding proportions for those supporting leave

4859/(4859+6371) ## Female

4479/(4746+4479) ## Male

## For Age

Data  %>%
  mutate(EU = ifelse(h_eumem == 1, "1", ifelse(h_eumem == 2, "0", NA))) %>%
  mutate(agegr = case_when(
                 between(h_age_dv, 16, 25) ~ "16 to 25 years",
                 between(h_age_dv, 26, 40) ~ "26 to 40 years",
                 between(h_age_dv, 41, 55) ~ "41 to 55 years",
                 between(h_age_dv, 56, 70) ~ "56 to 70 years",
                 h_age_dv > 70 ~ "Over 70 years"
        )) %>%
group_by(EU, agegr) %>%
summarise(n = n())

## Finding proportions for those supporting leave

736/(1749+736)

1508/(1508+2440)

2509/(2509+3013)

2737/(2646+2737)

1821/(1270+1821)


        ...
    
```


Write a couple of sentences interpreting your results.

Sex: There is a slightly higher proportion for support for leave for males when compared to females with a
difference of 5% between the groups. This is not signficant enough difference to identify a pattern of voting
with regards to the sex variable.

Age: The proportion of those supporting leave seems to increase as the you go from the younger age categories to
the higher age categories. For instance, the category of 16-25 had a 30% support for leave while the category of 
56-70 had 51% supporting leave. This implies a lower support for remain for older categories.
